Open-source tools in R for landscape ecology

### 2020-12-20 paper/paper.Rmd
improves Species distribution modeling

Species distribution modeling (SDM) analysis how landscape patterns (e.g. habitat suitability, resources availability, etc.) influence and determine the patterns of species distribution, mainly to infer ecological processes and predict future species distributions \cite{Wiersma2011}.
Originated in the 1970s, SDM has experienced numerous methodological advancement and a numerous body of literature exists today \cite{Zimmermann2010}.
Additionally, textbooks introducing basic concepts of SDM in *R* exists \cite{Guisan2017,Fletcher2019}.

Because the used modeling approaches are diverse \cite{Hooten2011,Kerr2011,Fletcher2019}, there is also a large number of *R* packages used for SDMs.
Popular approaches and packages include generalized linear models using, e.g., the *stats* \cite{RCoreTeam2019} package; generalized additive models using, e.g., the *mgcv* \cite{Wood2017} or *lme4* \cite{Bates2015} package; classification and regression trees (CART) using, e.g., the *rpart* \cite{Therneau2019}, *randomForest* \cite{Liaw2002} or *ranger* \cite{Wright2017} package or multivariate data analysis using , e.g., the *ade4* \cite{Dray2007} or *vegan* \cite{Oksanen2019} package.
Of course, also packages specifically designed for SDM exists.
This includes the *dismo* \cite{Hijmans2017}, *sdm* \cite{Naimi2016}, *ecospat* \cite{Broennimann2020}, *biomod2* \cite{Thuiller2020}, *PresenceAbsence* \cite{Freeman2008} or *zoon* \cite{Golding2018} packages.
Related to SDM, there are several packages to estimate home ranges and habitat selection, including *adehabitatHR* and *adehabitatHS* \cite{Calenge2006} or *amt \cite{Signer2019}*

>>>>

Species distribution modeling (SDM) examine how landscape patterns (e.g., habitat suitability or resources availability) influence and determine the patterns of species distribution, mainly to infer ecological processes and predict future species distributions \cite{Wiersma2011}.
Originated in the 1970s, SDM has experienced numerous methodological advancement, and a numerous body of literature exists today \cite{Zimmermann2010}.
Additionally, textbooks introducing basic concepts of SDM in *R* exists \cite{Guisan2017,Fletcher2019}.

Because the used modeling approaches are diverse \cite{Hooten2011,Kerr2011,Fletcher2019}, there is also a large number of *R* packages used for SDMs.
Popular approaches and packages include generalized linear models using, e.g., the *stats* \cite{RCoreTeam2019} package; generalized additive models using, e.g., the *mgcv* \cite{Wood2017} or *lme4* \cite{Bates2015} package; classification and regression trees (CART) using, e.g., the *rpart* \cite{Therneau2019}, *randomForest* \cite{Liaw2002} or *ranger* \cite{Wright2017} package or multivariate data analysis using , e.g., the *ade4* \cite{Dray2007} or *vegan* \cite{Oksanen2019} package.
Also, packages specifically designed for SDM exists, including includes the *dismo* \cite{Hijmans2017}, *sdm* \cite{Naimi2016}, *ecospat* \cite{Broennimann2020}, *biomod2* \cite{Thuiller2020}, *PresenceAbsence* \cite{Freeman2008}, or *zoon* \cite{Golding2018} packages.
Related to SDM, there are several packages to estimate home ranges and habitat selection, including *adehabitatHR* and *adehabitatHS* \cite{Calenge2006} or *amt \cite{Signer2019}*
